---
title: "Ellmer setup"
---

#### Load tidyverse and a library called `ellmer` that allows R to interact with LLMs.
```{r}
library(tidyverse)
library(ellmer)
```


#### Next, we load our API key into our code, replacing the text below with your copied API key:
```{r}
Sys.setenv(GROQ_API_KEY = "YOUR GROQ API KEY HERE")
```

#### Now we are ready to set up the model to talk to us when we use a variable called `chat`. Then we test it, by putting what we want to say to the LLM between quotes and parentheses.

```{r groq-setup}

# Set the model
chat <- chat_groq(
  model = "meta-llama/llama-4-scout-17b-16e-instruct"
)

chat$chat("Give me 20 names for a pet turtle")

```

#### Let's repeat the example using John Donne's poem, "No Man Is an Island":

```{r groq-poem}

poem <- "
No man is an island,
Entire of itself;
Every man is a piece of the continent,
A part of the main.
If a clod be washed away by the sea,
Europe is the less,
As well as if a promontory were:
As well as if a manor of thy friend's
Or of thine own were.
Any man's death diminishes me,
Because I am involved in mankind.
And therefore never send to know for whom the bell tolls;
It tolls for thee.
"

chat$chat(paste0("Extract all of the geographic terms in this poem: ", poem))

```
